Active Object Search with a Mobile Device for the Visually Impaired

### 2018-11-22 paper.tex
small typo revisions in sec1+2

\abstract{Modern smartphones can provide a multitude of services to assist people with visual impairments. The cameras embedded in these mobile devices, in particular, can be used for many useful tasks, like reading signs or searching for objects in unknown environments. Although previous research in computer vision has looked at ways to solve these tasks by processing the video frames of the mobile camera, very little work has been done in {\em actively} guiding the user towards specific points of interest that maximise the effectiveness of the underlying visual algorithms.
In this paper, we propose a control algorithm based on a Markov Decision Process (MDP) that generates instructions, in real-time, to guide a user searching for a target object with a smartphone's camera. The solution is part of a more general active vision application for the visually impaired. An initial implementation of the system on an Android smartphone was experimentally evaluated with sighted participants to determine the performance of the control algorithm. The results show the effectiveness of our solution and its potential application to help the visually impaired find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more normal and independent lives. Modern improvements to mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem.
%The ActiVis project aims to contribute to this effort by delivering an indoor navigation system targeted towards enabling blind and visually impaired (VI) users to safely navigate towards their desired destination or
The work presented here is part of a project to assist the visually impaired (VI) navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017}, but replaces the typical electro-mechanical actuators of a moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}. Schematically, the actuators used in a typical control loop, shown in Figure~\ref{fig:control-loop}, are replaced by a human. %This project expands upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

@@

The goal of our active search system is to understand the user's surroundings and determine what the next best course of action is to reach the target object based on what is currently within view and what has been observed in the past. To this end, we implemented a smartphone's guidance system based on a Markov Decision Process (MDP)~\cite{bellman1957markovian} that generates, in real-time, an optimal set of instructions (policy) for the user to point to the target, depending on a previously-learned spatial distribution of known objects and on the current camera's view.  
%
This work includes three main contributions:

@@

Section~\ref{sec:previous-work} discusses other relevant work done in this field, followed by a general explanation of the active vision system for human guidance in Sections~\ref{sec:system-design}, and a detailed explanation of the human-control module in Section~\ref{sec:controller-design}. The experimental results are presented in Section~\ref{sec:experiments}, after which we conclude the paper and discuss future work in Section~\ref{sec:conclusion}.

@@

\noindent Assistive vision and guidance for the VI is a growing research area. In recent years, the increase in mobile processing power and computer vision improvements have led to research in the use of smartphone cameras to augment or enhance a user's vision and help them find objects or other points of interest. Earlier attempts at the problem involved placing special markers or barcodes around an environment, which the user then scans with a smartphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.~Braille or sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more practical with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects, when they are in the camera's view, and to guide the user to them using sonified instructions~\cite{schauerte2012assistive}. This type of systems is more flexible than the tag-based ones, but it has the same drawback of being {\em passive}, in the sense that it relies on having the object within the camera's view in the first place. Also, no clear performance metrics are reported in the previous paper. The VizWize system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker who then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not enhance user independence, since the VI person is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, possibly limiting its use in some poor-reception areas.

>>>>

\abstract{Modern smartphones can provide a multitude of services to assist people with visual impairments. The cameras embedded in these mobile devices, in particular, can be useful for many tasks, like reading signs or searching for objects in unknown environments. Although previous research in computer vision has looked at ways to solve these tasks by processing the video frames of the mobile camera, very little work has been done in {\em actively} guiding the user towards specific points of interest that maximise the effectiveness of the underlying visual algorithms. In this paper, we propose a control algorithm based on a Markov Decision Process (MDP) that generates real-time instructions to guide a user searching for a target object with a smartphone's camera. The solution is part of a more general active vision application for the visually impaired. An initial implementation of the system on an Android smartphone was experimentally evaluated with sighted participants to determine the performance of the control algorithm. The results show the effectiveness of our solution and its potential application to help the visually impaired find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more normal and independent lives. Modern improvements in mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem. The work presented here is part of a project to assist the visually impaired (VI) navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017}, but replaces the typical electro-mechanical actuators of a moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}. Schematically, the actuators used in a typical control loop, shown in Figure~\ref{fig:control-loop}, are replaced by a human. %This project expands upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

@@

The goal of our active search system is to understand the user's surroundings and determine what the next best course of action is to reach the target object based on what is currently within view and what has been observed in the past. To this end, we implemented a smartphone guidance system based on a Markov Decision Process (MDP)~\cite{bellman1957markovian} that generates, in real-time, a series of instructions for the user to point to the target, depending on a previously-learned spatial distribution of known objects and on the camera's current view.  
%
This work includes three main contributions:

@@

Section~\ref{sec:previous-work} discusses other relevant work done in this field, followed by a general explanation of the active vision system for human guidance in Section~\ref{sec:system-design}, and a detailed explanation of the human-control module in Section~\ref{sec:controller-design}. The experimental results are presented in Section~\ref{sec:experiments}, after which we conclude the paper and discuss future work in Section~\ref{sec:conclusion}.

@@

\noindent Assistive vision and guidance for the VI is a growing research area. In recent years, the increase in mobile processing power and computer vision improvements have led to research in the use of smartphone cameras to augment or enhance a user's vision and help them find objects or other points of interest. Earlier attempts at the problem involved placing special markers or barcodes around an environment, which the user then scans with a smartphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.\ Braille or sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more practical with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects, when they are in the camera's view, and to guide the user to them using sonified instructions~\cite{schauerte2012assistive}. These type of systems is more flexible than the tag-based ones, but it has the same drawback of being {\em passive}, in the sense that it relies on having the object within the camera's view in the first place. Also, no clear performance metrics are reported in the previous paper. The VizWiz system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker who then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not enhance user independence, since the VI person is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, possibly limiting its use in some poor-reception areas.
